---
title: "dataAnalysis"
author: "Dany Chauvin"
date: "12/14/2020"
output: html_document
---

This is supposed to be a generic script for mother machine data analysis in Rstudio.

# Importing packages, functions and data

Restart R at this point.

```{r}
path_to_data_summary <- "/scicore/home/nimwegen/rocasu25/Documents/Projects/biozentrum/mothermachine_scripts/importingDataFromMomaToR/constitexprData.csv"
#Calling functions and packages
source("/scicore/home/nimwegen/rocasu25/Documents/Projects/biozentrum/mothermachine_scripts/importingDataFromMomaToR/loadFunctionsAndPackages.R")
#Import data
source("/scicore/home/nimwegen/rocasu25/Documents/Projects/biozentrum/mothermachine_scripts/importingDataFromMomaToR/deepMomaToDf.R")
```
# Setting strains

In multiplex experiments, different growth lanes might have different strains, characterized by different fluorescence backgrounds. The identity of these strains have to be attached to myframes_to_mycells and mycells. This can be attached to the data using .csv files or directly here.

```{r}
myframes_to_mycells <- myframes_to_mycells %>% 
   mutate(strain='asc662')
```
# Converting GFP units

Depending on the strain, fluorescence signal has to be converted to GFP numbers with different conversion factors.

```{r}
myframes_to_mycells <- myframes_to_mycells %>%
  # append relevant conversion parameters
  mutate(autofluo_predict=NA, fp_per_dn=NA,
         # case of GFPmut2 (from zaslaver library)
         autofluo_predict = ifelse(strain!='asc662', 133.6, autofluo_predict),
         fp_per_dn = ifelse(strain!='asc662', 0.198, fp_per_dn),
         # case of asc662
         autofluo_predict = ifelse(strain=='asc662', 422.8, autofluo_predict),
         fp_per_dn = ifelse(strain=='asc662', 0.0361 * 4, fp_per_dn)) %>% 
  # convert to gfp units (after subtracting autofluorescence)
  mutate(fluogfp_amplitude = fluo_amplitude - autofluo_predict * length_um,
         gfp_nb = fluogfp_amplitude * fp_per_dn ) %>% 
  group_by(cell) %>% 
  arrange(time_sec) %>% 
  mutate(g_birth=first(gfp_nb)) %>% 
  mutate(g_div=last(gfp_nb))

cell_fluo_info <- myframes_to_mycells %>% 
  distinct(cell,g_birth,g_div)
```

# Propagating fluorescence information to full cell cycle

```{r}
mycells <- mycells %>% 
  left_join(cell_fluo_info,by=c("cell")) %>% 
  mutate(c_birth=g_birth/l_birth,
         c_div=g_div/l_div,
         dg=g_div - g_birth,
         dcdt=(g_div/l_div - g_birth/l_birth) / div_time,
         g=log(g_div/g_birth) / div_time, # would be cleaner from a fit but difficult to compare
         gamma=dg/dl,
         q=dg/dl*alpha)
```

# Analysis

Now analysis can be performed. 


